Railway: Adaptive Disk Storage for Interaction Graphs

### 2014-09-11 docs/edbt2015/text/introduction.tex
some more progress

The ability to efficiently support historical analysis over interaction graphs
require effective solutions for the problem of \emph{data layout} on the disk.

% Importance of the data layout on disk and locality 
% Earlier work on blocks formed of temporal neighborlists

@@

@@

Typically, an interaction graph database is organized as blocks of temporal
neighbor lists.  Attributes can be stored in two ways, either separately
(e.g., in a relational table), or locally with the temporal neighbor lists. 
If they are stored separately, then the graph database cannot take advantage
of locality optimizations performed for block organization.  As discussed in
Gedik et al.~\cite{gedik14}, the database must go back and forth between the
disk blocks to access the edge attributes.  On the other hand, if attributes
are stored locally in the disk blocks containing the graph structure, then
there can be significant overhead due to disk I/O if only a few attributes are
needed to answer a query.

>>>>

The ability to efficiently support historical analysis over interaction graphs
require effective solutions for the problem of \emph{data layout} on disk.
Most graph algorithms are characterized by locality of
access~\cite{localityLayout}, which has been taken advantage of co-locating
edges in close proximity within the same disk blocks~\cite{gstoreThesis}. This
ensures that once a disk block is loaded from disk into the main memory
buffers, several edges can be used. The locality of reference is a result of
the traversal-based nature of most of the graph algorithms.

In interaction graphs, the locality of access is even more pronounced. First, 
the analysis to be performed on the interactions can be restricted to a
temporal view of the graph, such as finding the influential users over a given
week of interactions. This means that edges that are temporally close are to
be accessed together. Second, traversals are again key to many algorithms,
such as connected components, clustering coefficient, PageRank, etc. This
means that edges that are close by in terms of the path between their incident
vertices as well as their timestamps are to be located together with the same
blocks.

In our earlier work, we introduced an interaction graph database~\cite{gedik14}
that works on this principle of access locality. It proposed a disk
organization that consists of as set of blocks, each containing a list of
\emph{temporal neighbor lists}. A temporal neighborlist contains a head vertex
and a set of incident edges within a given time range. The layout optimizer
aims at bringing together temporal neighborlists that are (i) close in terms
of their temporal ranges, (ii) have many edges between them, and (iii) have
few edges going into temporal neighborlists outside the block.

@@

Attributes can be stored in two ways, either separately
(e.g., in a relational table), or locally with the temporal neighbor lists. 
If they are stored separately, then the graph database cannot take advantage
of locality optimizations performed for block organization.  As discussed in
Gedik et al.~\cite{gedik14}, the database must go back and forth between the
disk blocks to access the edge attributes.  On the other hand, if attributes
are stored locally in the disk blocks containing the graph structure, then
there can be significant overhead due to disk I/O if only a few attributes are
needed to answer a query.

@@
